RProtoBuf: Efficient Cross-Language Data Serialization in R

### 2014-01-03 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Improve introductory text.  Compare with the binary JSON formats like MessagePack and BSON.  Add the new high level figure about serializing/deserializing data from an interactive R session to other distributed systems.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@696 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Modern data collection and analysis pipelines are increasingly being
built using collections of components to better manage software
complexity through reusability, modularity, and fault
isolation \citep{Wegiel:2010:CTT:1932682.1869479}.  Different
programming languages are often used for the different phases of data
analysis -- collection, cleaning, analysis, post-processing, and
presentation in order to take advantage of the unique combination of
performance, speed of development, and library support offered by
different environments.  Each stage of the data
analysis pipeline may involve storing intermediate results in a
file or sending them over the network.  Programming languages such as
Java, Ruby, Python, and R include built-in serialization support, but
these formats are tied to the specific programming language in use.
% TODO(ms): and they often don't support versioning among other faults.
CSV files can be read and written by many applications and so are
often used for exporting tabular data.  However, CSV files have a
number of disadvantages, such as a limitation of exporting only
tabular datasets, lack of type-safety, inefficient text representation
and parsing, and ambiguities in the format involving special
characters.  JSON is another widely-supported format used mostly on
the web that removes many of these disadvantages, but it too suffers
from being too slow to parse and also does not provide strong typing
between integers and floating point.  Large numbers of JSON messages
would also be required to duplicate the field names with each message.

TODO(ms): Also work in reference to Split-Apply-Combine pattern for
data analysis \citep{wickham2011split}, since that is a great pattern
but it seems overly optimistic to expect all of those phases to always
be done in the same language.

@@

@@

\section{Protocol Buffers}

@@

>>>>

Modern data collection and analysis pipelines are increasingly being
built using collections of components to better manage software
complexity through reusability, modularity, and fault
isolation \citep{Wegiel:2010:CTT:1932682.1869479}.  
Data analysis patterns such as Split-Apply-Combine
\citep{wickham2011split} explicitly break up large problems into
manageable pieces.  These patterns are frequently employed with
different programming languages used for the different phases of data
analysis -- collection, cleaning, analysis, post-processing, and
presentation in order to take advantage of the unique combination of
performance, speed of development, and library support offered by
different environments.  Each stage of the data
analysis pipeline may involve storing intermediate results in a
file or sending them over the network.

Programming languages such as Java, Ruby, Python, and R include
built-in serialization support, but these formats are tied to the
specific programming language in use and thus lock the user into a
single environment.
%
% do not facilitate
% TODO(ms): and they often don't support versioning among other faults.
CSV files can be read and written by many applications and so are
often used for exporting tabular data.  However, CSV files have a
number of disadvantages, such as a limitation of exporting only
tabular datasets, lack of type-safety, inefficient text representation
and parsing, and ambiguities in the format involving special
characters.  JSON is another widely-supported format used mostly on
the web that removes many of these disadvantages, but it too suffers
from being too slow to parse and also does not provide strong typing
between integers and floating point.  Because the schema information
is not kept separately, multiple JSON messages of the same
type needlessly duplicate the field names with each message.
%
%
%
A number of binary formats based on JSON have been proposed that
reduce the parsing cost and improve the efficiency.  MessagePack
\citep{msgpackR} and BSON \citep{rmongodb} both have R interfaces, but
these formats lack a separate schema for the serialized data and thus
still duplicate field names with each message sent over the network or
stored in a file.  Such formats also lack support for versioning when
data storage needs evolve over time.

% TODO(mstokely): Take a more conversational tone here asking
% questions and motivating protocol buffers?

@@

\section{Protocol Buffers}
% This content is good.  Maybe use and cite?
% http://martin.kleppmann.com/2012/12/05/schema-evolution-in-avro-protocol-buffers-thrift.html

Protocol Buffers are a widely used modern language-neutral,
platform-neutral, extensible mechanism for sharing structured data.

one of the more popular examples of the modern

@@

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\textwidth]{protobuf-distributed-system-crop.pdf}
\end{center}
\caption{Example protobuf usage}
\label{fig:protobuf-distributed-usecase}
\end{figure}
